---
output: github_document
---

# yspec


## About

- Write data set specification data in `yaml` format
    - [About the format](./inst/doc/reference.md)
- The `yspec` package reads the specification and 
further formats the data
    - For example, [splitting data columns by other data items](./inst/doc/reference.md#split-data-column)
- Information can be accessed or utilized by other functions
in the ecosystem
- `yspec` will render a [`Define.pdf`](./inst/doc/define.pdf) document or output
in any other format supported by `rmarkdown`

## Use
```{r}
library(yspec)
```

We can load [this yaml file](inst/spec/spec.yml)

```{r}
sp <- load_spec("inst/spec/spec.yml")
```

See [here](./inst/doc/reference.md) for `YAML` specification details


A lot like the current spec setup
```{r,comment='.'}
names(sp)
sp[WT]$unit
```

```{r, results = "asis"}
yspec:::pander_table(sp)
```



Render in pdf format
```{r,echo=TRUE, results="hide", message=FALSE}
render_spec(sp, "define", output_dir="inst/doc")
```

with output [here](./inst/doc/define.pdf)


Or render to any format
```{r,echo=TRUE, results="hide", message=FALSE}
render_spec(sp, "define", output_format="md_document", output_dir="inst/doc")
```

Then we get [this output](./inst/doc/define.md)



## Build a `define.pdf` document

To make a proper `define.pdf` document, create another 
yaml file that points to data set specific files. 

For example [./inst/spec/project.yml](./inst/spec/project.yml)

Then render it:

```{r, echo = TRUE, results = "hide", message = FALSE}
render_define(file = "./inst/spec/project.yml", 
              output = "project", 
              title = "DEM104101 PK/PD Analysis Data Sets")
```

Output [here](./inst/doc/project.pdf)







