---
title: "Extract data from a yspec object"
author: ""
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_vignette: 
    keep_md: true
vignette: >
  %\VignetteIndexEntry{Extract data from a yspec object}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# Introduction
This vignette shows you how to extract information from a yspec object.

# Set up

```{r setup}
knitr::opts_chunk$set(comment = '.')
```


```{r, warning = FALSE, message = FALSE}
library(tidyverse)
library(yspec)
```

# Basic

First, recall that the yspec object is just a list.  We can use map to get 
information out of that list.  For example, to find the types of all the columns
in the data set use `purrr:::map`

```{r}
spec <- ys_help$spec()

map(spec, "type")[1:5]
```

If we wanted some column information in csv format we could also do

```{r}
map(spec, ~ paste(.x[["col"]],.x[["short"]],.x[["type"]],sep=','))[1:3]
```

There are other helper functions that do more than just pull a field.  We can 
extract a label and form a unit.

# Get units

Call `ys_get_unit` to get a named list of units. When the `parens` argument is
true, it will put parentheses around the unit.  Notice that the output here is 
different than if we were to only map across the list and ask for the unit field. 
Many columns don't have units and we'd get `NULL` back if we just mapped.  The 
`ys_get_unit` function subs in blank character data items when `unit` is `NULL`.

```{r}
ys_get_unit(spec, parens = TRUE) %>% unlist
```

# Get labels

We can also get labels

```{r}
ys_get_label(spec)[1:3]
```

# Get short

We can also get the short name

```{r}
ys_get_short(spec)[1:3]
```

# Get short with unit

This function returns a string with both the `unit` and `short`
```{r}
ys_get_short_unit(spec, parens = TRUE)[1:5] %>% as.list
```

# Meta information

The information you entered in `SETUP__` as well as other meta information is
stored as a list in the `meta` attribute of the `yspec` list. It can be 
accessed with `get_meta`.  This returns a list with the various pieces of meta 
data.

```{r}
get_meta(spec)[1:3]
```

You an pull a single piece of meta data with
```{r}
pull_meta(spec, "projectnumber")
```

## What is in the meta data?


```{r}
m <- get_meta(spec)
```

```{r}
names(m)
```

Some descriptions: 

1. `spec_file`: the yaml source file for the specification information
1. `spec_path`: the directory where the yaml source file is located
1. `name`: the stem of the yaml source file
1. `data_stem`: the stem that is used and it is assumed that data sets will be 
   written using this stem (for example `<data_stem>.csv` etc).
1. `data_path`: the assumed derived data directory

See also [reference.html](reference.html) for descriptions of meta fields.
