---
title: "Label the columns in a data set"
author: ""
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_vignette: 
    keep_md: true
vignette: >
  %\VignetteIndexEntry{Label the columns in a data set}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
# Introduction
This vignette shows you how to add labels to the columns of a data set.

# Set up

```{r setup}
knitr::opts_chunk$set(comment = '.')
```


```{r, warning = FALSE, message = FALSE}
library(tidyverse)
library(yspec)
```

# Load specification object and data set

We'll use the examples provided in the package

```{r}
data <- ys_help$data()
spec <- ys_help$spec()
```

The data
```{r}
as_tibble(data)
```

The spec
```{r}
spec
```


# Use ys_add_labels

```{r}
data <- ys_add_labels(data,spec)
```

It isn't obvious that anything was done here
```{r}
as_tibble(data)
```

How can you tell that the labels were added?

```{r}
labs <- map(data, attr, "label")

labs[1:5]
```

Or do this

```{r,eval=FALSE}
str(data)
```

# Where does label come from?

Ideally, we'd like to be writing in a `label` entry for every column in the 
data set.  You can set the `ys.require.label` option to `TRUE` to require this
when loading the spec (an error will be generated). 

But yspec has a function called `ys_get_label` that will form a label for you. 
Here are the rules:

1. If `label` exists for a column, it will be used
1. Otherwise, if `long` is found and it is `<=` 40 characters, it be used
1. Otherwise, `short` will be used; reminder that `short` defaults to the 
   column name (`col`) too

Let's look at some examples:

```{r}
ys_get_label(spec)[1:3]
```

```{r}
ys_get_label(spec$NUM)
```

```{r}
spec$NUM$label
```

```{r}
spec$C$label
```

# Custom label formation

Just as an example, we can add a custom labeling function. For example, I want 
the label to be the column name. 

Set up a function that takes the column data as the first argument
```{r}
label_fun <- function(x,...) x[["col"]]
```

Now, pass that function into `ys_add_labels`

```{r}
data <- ys_add_labels(data, spec, fun = label_fun)
```

And check the output

```{r}
map(data, attr, "label")[1:5]
```

# Extract the label field

Recall that the yspec object is just a list.  We can always map across that list
and grab the label field:s

```{r}
map(spec, "label")[1:5]
```

