---
title: "Project-wide column definitions"
author: ""
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_vignette: 
    keep_md: true
vignette: >
  %\VignetteIndexEntry{yspec}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# Introduction
This document shows how data set column definitions can be entered into 
a lookup file which can be accessed by multiple data specification files 
within a project. This document also discusses an internal lookup data base
that is always available for individual data sets to look up standardized 
column information for commonly used data items in our workflow.

# Create two files

## The lookup file

The lookup file that is to be accessed by other data specification files
is just another data specification file.  For example, create a file called
`lookup.yml` and enter this information.

```yaml
# in file lookup.yml
AMT: 
  short: dose amount
  unit: nM
  type: numeric
WT: 
  short: patient weight
  unit: lbs
```

This information must be valid yspec data specification format and (generally)
valid yaml. 

## The specification file

This is just the standard data specification file

```yaml
SETUP__:
  description: PKPD analysis data set
  lookup_file: lookup.yml
C: 
  short: comment character
AMT: !look
```
Notice two things about this file: we included a `lookup_file` section in the 
`SETUP__` section and we referenced our `lookup.yml` file.  By default, 
yspec expects that the lookup file is in the same directory as the 
spec file.  Also, in the `AMT` column, we used the `!look` handler to 
indicate that we wanted that data to be looked up. 

Alternatively, we could just pass in empty data and yspec will assume that 
you want to try to look up that data

```yaml
SETUP__:
  description: PKPD analysis data set
  lookup_file: lookup.yml
C: 
  short: comment character
AMT:
```


# Internal lookup data base

There is an internal data base of common data set columns that 
yspec will attach by default.  So, with no lookup file defined, 
we could write the following in our specification file

```yaml
SETUP__:
  description: PKPD analysis data set
  use_internal_db: true
C: 
AMT:
MDV:
EVID: 
WT:
EGFR:
ALB: !look
ZIP_CODE: 
  values: 55378
```


```{r,echo = FALSE}
code <- '
SETUP__:
  description: PKPD analysis data set
  use_internal_db: true
C: 
AMT:
MDV:
EVID: 
WT:
EGFR:
ALB: !look
ZIP_CODE: 
  values: 55378
'

writeLines(code, file <- file.path(tempdir(), 'spec.yml'))
```

We can read this data in and have the columns defined

```{r,message=FALSE}
library(yspec)
library(dplyr)

spec <- ys_load(file)

spec
```


# Tracking the lookup status of each column

This all can get confusing about where each column is coming from.  You can 
audit the spec object and find you where a lookup event happened

```{r}
ys_lookup_source(spec)
```


Here, we can see that most of the columns came from the internal data base and 
that the one column (`ZIP_CODE`) came by our own specification. 

You can also re-create the lookup object (just a named list) for a 
specification object.  Just click open the arrow to see the output.


<details>

```{r}
ys_get_lookup(spec) %>% glimpse()
```

</details>


















