---
title: "Project Object and Define Document"
author: "Kyle Baron"
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_vignette: 
    keep_md: true
vignette: >
  %\VignetteIndexEntry{Project Object and Define Document}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
# Introduction
This vignette shows you how to gather the specifications for all of your
analysis project data sets, assemble into a single object, and use that 
object to render a project-wide define document.

# Set up

```{r}
library(yspec)
```

# Load specifiation objects for all data sets

If my project had 3 analysis data sets, one for PK, 
one for PKPD, and one for AEs, we create a specification 
object for each one.  Here, we're using `load_spec_ex()`
because we're using example material from the package.  
You should use `load_spec()` for your project.

```{r}
sp1 <- load_spec_ex("DEM104101F_PK.yml")
sp2 <- load_spec_ex("DEM104101F_PKPD.yml")
sp3 <- load_spec_ex("DEM104101F_AE.yml")

```

So each object looks something like
```{r}
sp1
```

So this includes our usual information about the columns in the 
data set. 


# Create a project object

We take these specification objects and create a 
project object from them. 
```{r}
proj <- as_proj_spec(sp1,sp2,sp3)

proj

```

Now, for each specification object, we have a data / specification
name as well as a description (along with some other information). 

This file is by default saved to a temporary location.  But you 
can also save it locally and keep it as a part of the project.

```{yaml, code = readLines(get_meta(proj)[["yml_file"]])}

```


To do this, we tucked some additional information into the 
data set specification file
```{r}
str(get_meta(sp1))
```

In the `SETUP__` block, we included a description; this is 
required to make a project object from data objects. There also 
is a `name` that defaults to the stem of the data specification 
file.  We also optionally included `projectnumber` and 
`client`.  These will get inserted into the define document
when the MetrumRG Rmarkdown template is used.


# Render a define file
```{r, eval=FALSE}
render_define(proj, stem = "define_doc")
```





