---
title: "Project object and define document"
author: ""
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_vignette: 
    keep_md: true
vignette: >
  %\VignetteIndexEntry{Project Object and Define Document}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
# Introduction
This vignette shows you how to gather the specifications for all of your
analysis project data sets, assemble into a single object, and use that 
object to render a project-wide define document.

# Set up

```{r}
library(yspec)
```

# Load specification objects for all data sets

If my project had 3 analysis data sets, one for PK, one for PKPD, and one for 
AEs, we create a specification object for each one.  Here, we're using 
`load_spec_ex()` because we're using example material from the package.  You 
should use `ys_load()` for your project and give the full path to the 
yaml specification file. 

```{r}
sp1 <- load_spec_ex("DEM104101F_PK.yml")
sp2 <- load_spec_ex("DEM104101F_PKPD.yml")
sp3 <- load_spec_ex("DEM104101F_AE.yml")

```

Each object looks something like
```{r}
sp1
```

This includes our usual information about the columns in the data set. 


# Create a project object

We take these specification objects and create a 
project object from them. 
```{r}
proj <- ys_project(sp1,sp2,sp3)

proj

```

Now, for each specification object, we have a data / specification name as well 
as a description (along with some other information). 

This file is by default saved to a temporary location.  But you can also save 
it locally and keep it as a part of the project.

```{yaml, code = readLines(get_meta(proj)[["spec_file"]])}

```


To do this, we tucked some additional information into the data set 
specification file. For example, the specification file for the first spec 
object (`sp1`) looks like this:
```{yaml,code = readLines(yspec_yml_file(sp1))}

```

In the `YPROJ__` block, we included a description; this is required to make a 
project object from data objects. There also is a `name` that defaults to the 
stem of the data specification file.  We also optionally included `projectnumber` 
and `sponsor`.  These will get inserted into the define document when the 
MetrumRG Rmarkdown template is used.

# Render a define file
Now, we have an object with the project information put together.  The essential
contents of that object have been written to a project spec file and can 
be used to template the creation of a project data specification file.

```{r, eval=FALSE}
ys_document(proj, stem = "define_doc")
```





