---
title: "YAML specification"
author: ""
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{yspec yaml specification}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

## Syntax

yspec uses standard [yaml](https://en.wikipedia.org/wiki/YAML) syntax to state 
the data set column definitions.

__NOTES__

- Use double-quotes around any value you want to be a string
- `true` and `yes` by themselves will be rendered as `TRUE`; use `"yes"` if you 
  need that word by itself as a value for a field
- `false` and `no` by themselves will be returned as `FALSE`; use `"no"` if you
  need that word by itself as a value for a field
- If a string starts with a piece of punctuation, make sure to put the entire 
  string in double quotes
    - Use `short: "> QL"` not `short: > QL`
    - Use `values: [".", C]` not `values: [.,C]`
    - Use `address: "123 Main St."` not `address: 123 Main St.`

Instructions for including TeX in the yaml specification code are provided
in a section below. 

## Organization
Save your data specification code in a file, typically with a `.yaml` 
file extension.  

At the top of the file, include a block called `SETUP__:`; this is where 
the data set meta data is stored.  For example

```yaml
SETUP__:
  description: PKPD analysis data set. 
  use_internal_db: true
  projectnumber: FOO123
  sponsor: MetrumRG
```

See the details below for other files that can be included here.

Next, list each data set column in order, with the data column name 
starting in the first column and ending with a colon.  For example:

```yaml
WT:
  short: weight
  unit: kg
  range: [50, 150]
```

This specifies a "short" name for this column as well as a unit and a range. 
A complete listing is provided below. 

You can see an fully worked example by running

```{r, eval = FALSE}
ys_help$yaml()
```

See the `?ys_help` help topic for more information.

Or, you can export a collection of package assets with this command

```{r, eval = FALSE}
ys_help$export(output="assets")
```

See the [ys_help] topic for more information.  

## `SETUP__` specification fields

- `description`: `<character>`; a short description of the data set 
- `projectnumber`: `<character>` the project reference number; may be 
   incorporated into rendered define documents; when the project number is 
   given in the first yspec object in a project object, that project number 
   will be rendered in the project-wide define document
- `sponsor`: `<character>` the project sponsor; when the project sponsor is 
   given in the first yspec object in a project object, that project sponsor 
   name will be rendered in the project-wide define document 
- `lookup_file`: `<character>`; a yaml array of other yaml files where 
   yspec will look for column lookup information
- `use_internal_db`: `<logical> (true/false)`; if `true`, then yspec will 
   load the internal column lookup database
- `glue`: `<map>`; specify name/value pairs;  in the yaml data specification, 
   use `<<name>>` in the text and `value` will glued into the text after
   it has been sanitized; intended use is to allow LaTeX code to evade the 
   sanitizer

## Data column specification fields

- `short: short-name` 
    - a short name for the column; don't include unit here
- `unit: numeric` 
    - the unit of measure
- `range: [min-value, max-value]`
    - indicates continuous data
- `values: [val1, val2, valn]`
    - specify each valid value
    - indicates discrete data
- `values: {decode1: val1, decode22: val2}`
    - put the decode into the specification of values using a default 
      yaml map structure (the decode is to the left of the `:`)
    - __Note__ the curly brackets, not square brackets
- `decode: [decode1, decode2, decode3]`
    - Separate the `decode` from the `values` specification
    - See example below for clearer way to input very long decodes
- `longvalues: true`
    - print the values in a (long) `yaml`-formatted list
- `comment: just whatever you want to say`
- `comment: >
      say something
      on multiple lines of
      text`
- `source: ADSL.xpt`
    - Where the data came from
    - include both the sdtm domain and variable name
- `about: [short-name, unit]`
  - This is a convenience structure.
- `long: a longer name to describe the column`
- `dots`: 
  - A named list of whatever you want to carry along in the object; the `dots` list
    isn't used by any rendering function in the yspec package, but might be used 
    by a custom rendering function 
- `axis`:
  - A short-ish name that can be used for axis titles for plots
  - Generally, don't include unit; yspec helpers will add that 
    automatically by default
  - If `short` will work for your axis title (as it is ... with no 
    modification), yspec will use that if no `axis` field is used
- `type`:
  - Can be `numeric`, `character`, or `integer`
  - This is optional; the default is `numeric`
- `lookup`:
  - Logical; if `true` then the definition for the column is looked up in 
  the `lookup_files` (specified in `SETUP__:`)
  - Use the `!look` handler to indicate lookup
  

## Defaults

- If `type` is not given, then it will default to `numeric`

## Examples

### Continuous values

- The `about` array provides a short name and unit
- Any time `range` is given, the data is assumed to be continuous

```yaml
WT:
  about: [Weight, kg]
  range: [5,300]
```

This is equivalent to 

```yaml
WT:
  short: Weight
  unit: kg
  range: [5,300]
```

### Character data
- Using `values` indicates discrete data

```yaml
RACE:
  values: [white, black, Native American, Other]
```

Any other array input structure can be used.  For example

```yaml
RACE: 
  values:
    - white
    - black
    - Native American
    - Other
```


By default, `values` are printed as comma-separated list.  To
get them to print in long format

```yaml
RACE:
  values: [white, black, Native American, Other]
  longvalues: true
```

### Discrete data with decode

__Method 1__

- Notice that the yaml key can only be simple character value
- Also, we use curly braces to specify a list like this
- Finally it is a `:` that separates decode (on the left) and the 
  value (on the right).

```yaml
SEX:
  values: {dude: 0, gal: 1}
```

Special handlers are available that add some flexibility to this value / 
decode specification.  

The `!value:decode` handler allows you to put the value on the left and 
decode on the right

```yaml
SEX: 
  values: !value:decode
    0 : dude
    1 : gal
```

The default behavior can be achieved with

```yaml
SEX: 
  values: !value:decode
    dude: 0
    gal: 1
```

The handlers also allow associating multiple values with a single decode

__To get multiple values with the same decode__

```yaml
STUDY:
  values: !decode:value
    phase 1 : [101, 102, 103]
    phase 2 : [201, 202, 203]
    phase 3 : [301, 302, 303]
```


__Method 2__

- These are more complicated decodes
- Put the `values` and `deode` in brackets (array)

```yaml
BQL:
  values: [0,1]
  decode: [not below quantitation limit, below quantitation limit]
```

__Method 3__
Really, it's the same as method 2, but easier to type and read when the
decode gets really long

```yaml
BQL:
  values [0, 1]
  decode:
    - not below the quantitation limit of 2 ng/ml
    - below the quantitation limit of 2 ng/ml
```

### Look up column definition

Either fill in the `lookup` field or use the `!look` handler

```yaml
CMT: 
  lookup: true
```

```yaml
CMT: !look
```
You can also give the column name to import

```yaml
HT: 
  lookup: HT_INCHES
```

In this example, there would be a column called `HT_INCHES` in the 
lookup file that would be imported under the name `HT`.


### Include TeX in data specification document

Most define documents get rendered via `xtable` and the text gets processed
by a sanitize function.  yspec implements a custom sanitize function
called `ys_sanitize`, which is similar to `xtable::sanitize`, but whitelists
some symbols so they do not get sanitized.  

To protect TeX code from the sanitizer, first create a field in `SETUP_`
called `glue` with a map between a name and some corresponding TeX code.  In 
the following example, we with to write $\mu$g/L, so we create a name 
called `mugL` and map it to `$\\mu$g/L`:

```{yaml}
SETUP__:
  glue: {mugL: "$\\mu$g/L"}
```

Once the map is in place, we can write the data set column definition like 
this:

```{yaml}
DV: 
  unit: "<<mugL>>"
```

When the table for the define document is rendered, first the sanitizer will 
run, but it won't find anything in the `unit` field for the `DV` column.  Then 
yspec will call `glue()` and replace `<<mugL>>` with `$\\mu%g/L`.  

Notice that we put all of the values in quotes;  this is good practice to ensure
that yaml will parse the value as a character data item when reading in the 
spec. 













