---
title: "Yaml specification reference"
author: ""
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_vignette: 
    toc: true
    number_sections: true
vignette: >
  %\VignetteIndexEntry{Yaml specification reference}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

# Syntax

yspec uses standard [yaml](https://en.wikipedia.org/wiki/YAML) syntax to state 
the data set column definitions.

__NOTES__

- Use double-quotes around any value you want to be a string
- `true` and `yes` by themselves will be rendered as `TRUE`; use `"yes"` if you 
  need that word by itself as a value for a field
- `false` and `no` by themselves will be returned as `FALSE`; use `"no"` if you
  need that word by itself as a value for a field
- If a string starts with a piece of punctuation, make sure to put the entire 
  string in double quotes
    - Use `short: "> QL"` not `short: > QL`
    - Use `values: [".", C]` not `values: [.,C]`
    - Use `address: "123 Main St."` not `address: 123 Main St.`
    - Use `label: "line 1 \\n line 2"` not `label: line 1 \n line2`

Instructions for including TeX in the yaml specification code are provided
in a section below. 

# Organization
Save your data specification code in a file, typically with a `.yaml` 
file extension.  

At the top of the file, include a block called `SETUP__:`; this is where 
the data set meta data is stored.  For example

```txt
SETUP__:
  description: PKPD analysis data set
  use_internal_db: true
  projectnumber: FOO123
  sponsor: MetrumRG
```

See the details below for other files that can be included here.

Next, list each data set column in order, with the data column name 
starting in the first column and ending with a colon.  For example:

```txt
WT:
  short: weight
  unit: kg
  range: [50, 150]
```

This specifies a "short" name for this column as well as a unit and a range. 
A complete listing is provided below. 

You can see an fully worked example by running

```{r, eval = FALSE}
ys_help$yaml()
```

See the `?ys_help` help topic for more information.

Or, you can export a collection of package assets with this command

```{r, eval = FALSE}
ys_help$export(output="assets")
```

See the [ys_help] topic for more information.  

# `SETUP__` specification fields

- `description`: `<chr>`; a short, label-like description of the data set 
- `projectnumber`: `<chr>` the project reference number; may be 
   incorporated into rendered define documents; when the project number is 
   given in the first yspec object in a project object, that project number 
   will be rendered in the project-wide define document
- `sponsor`: `<chr>` the project sponsor; when the project sponsor is 
   given in the first yspec object in a project object, that project sponsor 
   name will be rendered in the project-wide define document 
- `data_path`: `<chr>`; a path locating the data set associated with the spec
- `data_stem`: `<chr>`; the stem (no extension) for the data set associated 
  with the spec; usually the stem of the data file is the same as the stem of 
  the spec, but they can also be different
- `lookup_file`: `<chr>`; a yaml array of other yaml files where 
   yspec will look for column lookup information
- `extend_file`: `<chr>`; the name of a single yspec yaml file that can 
  be used to "extend" the current yspec object
- `use_internal_db`: `<logical> (true/false)`; if `true`, then yspec will 
   load the internal column lookup database
- `import`: `<chr>`;  give the name of a `<file>` to import into the 
  current data spec; all 
  columns from `<file>` are imported as is; additional columns may also be listed
  with the normal syntax and these columns will appear after the imported 
  columns
- `character_last`: `<logical>`;  if true, automatically push all non-numeric 
  columns to the end of the data specification list
- `comment_col`: `<chr>`; identify the column that is used to store 
  comments; the comment column will not be pushed to the back when 
  `character_last` is true
- `glue`: `<map>`; specify name/value pairs;  in the yaml data specification, 
  use `<<name>>` in the text and `value` will glued into the text after
  it has been sanitized; intended use is to allow LaTeX code to evade the 
  sanitizer
- `max_nchar_label`: `integer`; the maximum number of characters allowed in 
  the `label` field
- `max_nchar_short`: `integer`; the maximum number of characters allowed in 
  the `short` field
- `max_nchar_col`: `integer`; the maximum number of characters allowed in 
  the data set column name
- `flags`: `<map>`; for each key, an array of column names where a logical 
  data item will be set in the `dots` list

# Data column specification fields

- `short: short-name` 
    - a short name for the column; don't include unit here
    - this should be no more than 40 characters by default
    - use `max_nchar_short` to change the maximum, but this is not recommended
- `unit: numeric` 
    - the unit of measure
- `range: [min-value, max-value]`
    - indicates continuous data
- `values: [val1, val2, valn]`
    - specify each valid value
    - indicates discrete data
- `values: {decode1: val1, decode22: val2}`
    - put the decode into the specification of values using a default 
      yaml map structure (the decode is to the left of the `:`)
    - __Note__ the curly brackets, not square brackets
- `decode: [decode1, decode2, decode3]`
    - separate the `decode` from the `values` specification
    - see example below for clearer way to input very long decodes
- `longvalues: true`
    - print the values in a (long) `yaml`-formatted list
- `comment: just whatever you want to say`
- `comment: >
      say something
      on multiple lines of
      text`
- `source: ADSL.xpt`
    - where the data came from
    - include both the sdtm domain and variable name
- `about: [short-name, unit]`
  - this is a convenience structure
- `label: label-text`
  - a label for the column
  - the label must be 40 or fewer characters by default
  - use `max_nchar_label` to change the maximum, but this is not recommended
  - this will get written into the define file as well as the data frame prior to writing out to sas xport format
- `long: a longer name to describe the column`
- `dots`: 
  - a named list of whatever you want to carry along in the object; the `dots` 
    list isn't used by any rendering function in the yspec package, but might 
    be used by a custom rendering function 
- `axis`:
  - a short-ish name that can be used for axis titles for plots
  - generally, don't include unit; yspec helpers will add that 
    automatically by default
  - if `short` will work for your axis title (as it is ... with no 
    modification), yspec will use that if no `axis` field is used
- `type`:
  - can be `numeric`, `character`, or `integer`
  - this is optional; the default is `numeric`
- `make_factor`: if `true`, then the column will be able to be converted to a 
  factor regardless of whether `decode` is included or not
- `lookup`:
  - logical; if `true` then the definition for the column is looked up in 
  the `lookup_files` (specified in `SETUP__:`)
  - use the `!look` handler to indicate lookup
  
# Maximum character limits

There are character count maximums for some of the input fields. These limits
are in place because we frequently write datasets out in SAS-transport format
and there limits associated with that output file, depending on which version 
you are writing to.

All the default character counts can be overrided by options in `SETUP__:`; 
however the user is warned to think carefully before overriding the defaults, 
since these are in place for a reason. If any default is overridden, the user
takes responsibility for any downstream consequences, including the possibility
that the data set will fail to meet regulatory standards.

# Namespaces

Namespaces are alternative representation of certain column data fields 

- `unit`
- `short`
- `label`
- `long`
- `decode`
- `comment`

You can create namespaces by attaching a `.<name>` suffix to eligible fields. 

For example, we can create a "tex" representation for `unit` like this

```txt
DV: 
  short: dependent variable
  unit: "microgram/mL"
  unit.tex: "$\\mu$g/mL"
```

Here, the `unit:` entry states the value for unit in the `base` namespace, the 
default data you get on load. Using `unit.tex:` introduces an entry for the 
`tex` namespace. After loading the spec, you can change to this namespace using

```{r, eval = FALSE}
spec <- ys_load(...)
spec_tex <- ys_namespace(spec, "tex") 
```

Any time you attach a `.<name>` suffix to a field, `yspec` will interpret that 
as an attempt to enter namespace data. The user is responsible for creating 
and organizing namespaces and naming them. `yspec` will create the `base` 
namespace. Also, when rendering a data specification document, `yspec` will 
attempt to switch to the `tex` namespace if it exists. Beyond that, `yspec` is 
agnostic to the names of the namespaces you create.

As another example, we can have alternate `short` names depending on whether 
or not we are using that name to create axis titles for a plot

```txt
EGFR:
  short: estimated creatinine clearance
  short.plot: eGFR
```

or decode

```txt
SEX:
  values: [0, 1]
  decode: [male, female]
  decode.letter: [m, f]
```

# Defaults

- If `type` is not given, then it will default to `numeric`

## Examples

## Continuous values

- The `about` array provides a short name and unit
- Any time `range` is given, the data is assumed to be continuous

```txt
WT:
  about: [weight, kg]
  range: [5, 300]
```

This is equivalent to 

```txt
WT:
  short: weight
  unit: kg
  range: [5,300]
```

## Character data
- Using `values` indicates discrete data

```txt
RACE:
  values: [White, Black, Native American, Other]
```

Any other array input structure can be used.  For example

```txt
RACE: 
  values:
    - White
    - Black
    - Native American
    - Other
```


By default, `values` are printed as comma-separated list.  To
get them to print in long format

```txt
RACE:
  values: [White, Black, Native American, Other]
  longvalues: true
```

## Discrete data with decode

__Method 1__

- Notice that the yaml key can only be simple character value
- Also, we use curly braces to specify a list like this
- Finally it is a `:` that separates decode (on the left) and the 
  value (on the right).

```txt
SEX:
  values: {dude: 0, gal: 1}
```

Special handlers are available that add some flexibility to this value / 
decode specification.  

The `!value:decode` handler allows you to put the value on the left and 
decode on the right

```txt
SEX: 
  values: !value:decode
    0 : dude
    1 : gal
```

The default behavior can be achieved with

```txt
SEX: 
  values: !value:decode
    dude: 0
    gal: 1
```

The handlers also allow associating multiple values with a single decode

__To get multiple values with the same decode__

```txt
STUDY:
  values: !decode:value
    phase 1 : [101, 102, 103]
    phase 2 : [201, 202, 203]
    phase 3 : [301, 302, 303]
```


__Method 2__

- These are more complicated decodes
- Put the `values` and `deode` in brackets (array)

```yaml
BQL:
  values: [0,1]
  decode: [not below quantitation limit, below quantitation limit]
```

__Method 3__
Really, it's the same as method 2, but easier to type and read when the
decode gets really long

```txt
BQL:
  values [0, 1]
  decode:
    - not below the quantitation limit of 2 ng/ml
    - below the quantitation limit of 2 ng/ml
```

## Look up column definition

Either fill in the `lookup` field or use the `!look` handler

```txt
CMT: 
  lookup: true
```

```txt
CMT: !look
```
You can also give the column name to import

```txt
HT: 
  lookup: HT_INCHES
```

In this example, there would be a column called `HT_INCHES` in the 
lookup file that would be imported under the name `HT`.


## Include TeX in data specification document

Most define documents get rendered via `xtable` and the text gets processed
by a sanitize function.  yspec implements a custom sanitize function
called `ys_sanitize()`, which is similar to `xtable::sanitize`, but whitelists
some symbols so they do not get sanitized.  

To protect TeX code from the sanitizer, first create a field in `SETUP__`
called `glue` with a map between a name and some corresponding TeX code.  In 
the following example, we with to write $\mu$g/L, so we create a name 
called `mugL` and map it to `$\\mu$g/L`:

```{txt}
SETUP__:
  glue: {mugL: "$\\mu$g/L"}
```

Once the map is in place, we can write the data set column definition like 
this:

```{txt}
DV: 
  unit: "<<mugL>>"
```

When the table for the define document is rendered, first the sanitizer will 
run, but it won't find anything in the `unit` field for the `DV` column.  Then 
yspec will call `glue()` and replace `<<mugL>>` with `$\\mu%g/L`.  

Notice that we put all of the values in quotes;  this is good practice to ensure
that yaml will parse the value as a character data item when reading in the 
spec. 

## `flags`

The `flags` section in `SETUP__:` is available for you to name sets of columns
in the work in spec. For example, the following code defines a flag called 
`covariate` and it names three columns (`WT`, `AGE`, and `CRCL`) to carry 
this tag

```{txt}
SETUP__:
  flags:
    covariate: [WT, AGE, CRCL]
```

When yspec loads a yaml file that contains `flags`, it will go into every column
in the spec and add a logical flag in `dots` that indicates whether or not that 
column is a member of that covariate set. For this example, all columns in the 
spec will have `dots$covariate` set to `FALSE` except for `WT`, `AGE`, and 
`CRCL` where it will be set to `TRUE`. 

The user can appear to this information when filtering the spec. Filtering like
this will return a yspec object containing only `WT`, `AGE`, and `CRCL`. 

```{r, eval = FALSE}
ys_filter(spec, covariate)
```

Note that this flagging process will not overwrite a flag that the user already
set in a specific column. In this example, `AGE` will not be flagged as a 
covariate, but `WT` and `CRCL` will.  

```{txt}
SETUP__:
  flags:
    covariate: [WT, AGE, CRCL]
WT: 
  short: weight
AGE: 
  short: age
  dots: {covariate: false}
CRCL:
  short: creatinine clearance
```

It's recommended that `flags` are given in the `SETUP__` information only, but 
the user can override as needed. 

