---
title: "yspec"
author: "Kyle Baron"
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_vignette: 
    keep_md: true
vignette: >
  %\VignetteIndexEntry{yspec}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---



## About

The `yspec` package will read you data specification file
when it is written in a specific `yaml` format.  

```{r "library"}
library(yspec)
```

## Example spec file

An example specification file looks like:

```{r "specfile", comment = " "}
specfile <- yspec::file_spec_ex()

cat(readLines(specfile)[1:29], sep = "\n")
```


Once the data specification `yaml` file is written, 
it can be loaded in `R`

```{r "load_specfile", comment = " "}
spec <- load_spec(specfile)

spec
```

Data from specific columns can be printed
```{r "column", comment = " "}
spec$WT
```

or summarized
```{r "column_summary", comment = ' '}
summary(spec, WT, DV, EGFR)
```

## Example to render spec

The specification object can be rendered to a specification file
```{r "render_spec", results = "hide", message = FALSE}
render_spec(spec, stem = "render_spec")
```

With output [here](render_spec.pdf).


## Example project file

To create a project-wide data specification document, first
create a project file:
```{r "proj_file", comment = " "}
proj <- file_proj_ex()

cat(readLines(proj), sep = "\n")
```

The project file refers to other data specification documents, all of 
which can be rendered to a single comprehensive 
data specication file.


## Render a project file

```{r "fda_define", results = "hide", message = FALSE}
render_fda_define(proj)
```

With output [here](define.pdf).
