---
title: "Get started"
author: ""
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_vignette: 
    keep_md: true
vignette: >
  %\VignetteIndexEntry{Get started}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---



## About

The `yspec` package will read your data specification file when it is written 
in a specific `yaml` format. Using the object created from this file, you can 
validate data assembly outputs, label the data frame to be written in sas xport
format, create a `define.pdf` document, and more.

For each data set in your project that needs documentation, create
a yaml file that lists the columns in the data set along with details 
about the data in that column.  This yaml file can be loaded into your 
R session into an object that you can work with. This is referred to as
a `spec` object.  The term `spec` refers to a single documented data object / 
data file. 

Once all of your data sets have been documented with their own yaml file, 
you can create another object called a `yproj` objects.  This is used to 
template the rendering of a single integrated data definitions file for 
your entire project. 

Keep reading the vignette to see how it works.

<hr>

```{r "library"}
library(yspec)
```

## Example spec file

An example specification file looks like:

```{r "specfile", comment = " "}
specfile <- ys_help$file()

cat(readLines(specfile)[1:29], sep = "\n")
```


Once the data specification `yaml` file is written, 
it can be loaded in `R`

```{r "load_specfile", comment = " "}
spec <- ys_load(specfile)

spec
```

Data from specific columns can be printed
```{r "column", comment = " "}
spec$WT
```

or summarized
```{r "column_summary", comment = ' ', out.width = 120}
summary(spec, WT, DV, EGFR)
```

### Items that you should be including for most columns

- `short`: a short name for the column (e.g `weight`); this will default to 
  the column name (`col`) ... sometimes that makes sense
- `label`: to be used to label the data set and to populate the `define.pdf` 
  document (e.g. `patient weight at baseline`)
- `unit`: when it's appropriate (e.g. `kg`)
- `decode`: for discrete data items (e.g. if SEX is `values: [0,1]` then then 
  use `decode: [male, female]`; or include the `make_factor: true` field


## Check a data set against the spec

Use the `ys_check()` function, with the data frame as the first argument 
and the spec object as the second argument

```{r}

data <- ys_help$data()

ys_check(data, spec)

```


## Example to render spec

The specification object can be rendered to a specification file with the 
`ys_document` function

```{r "render_spec", results = "hide", message = FALSE, warning = FALSE}
ys_document(spec, stem = "working_document")
```

With output [here](working_document.pdf).

`ys_document` will pass along arguments to `rmarkdown::render` so that you 
can control those aspects of how the document is rendered. You can also 
create custom output formats to get the data table to render in the way that you
like.

## Example project object

To create an project-wide listing of documented data sets, we create a 
`yproj` or project object.  We create this from the spec objects that we 
read about in the previous section.  Let's load another object to use along 
with the object loaded in the previous section.

```{r}
pdspec <- load_spec_ex("DEM104101F_PKPD.yml")
```

Now, we have two objects to work with:

```{r}
head(spec)

head(pdspec)
```

We can create a project object from both objects
```{r}
proj <- ys_project(spec,pdspec)

proj
```



## Render a project file

### Working document

To render the project file we'll use the same `ys_document()` function.  
This time, we'll add some extra (optional) arguments that will help us 
get the document to look the way we want:

```{r "render_define", results = "hide", message = FALSE, warning = FALSE}
ys_document(
  proj, 
  stem = "project_document", 
  build_dir = mrgtemplate(),
  author = "Michelle Johnson", 
  title = "Analysis data specification"
)
```

With output [here](project_document.pdf).

Using the `build_dir` argument gets us the document rendered with Metrum 
Research Group branding.  Also, `author` and `title` are passed into the
configuration fields for this document.

### Regulatory document 

To get a document that is formatted according to FDA requirements, use:

```{r "render_fda", results = "hide", message = FALSE, warning = FALSE}
ys_document(
  proj, 
  type = "regulatory",
  stem = "fda_document", 
  build_dir = mrgtemplate(),
  author = "Michelle Johnson", 
  title = "Analysis data specification"
)
```

With output [here](fda_document.pdf).
